---
title: "Data Visualization Portfolio"
author: "Edward Visel"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: "lumen"
    highlight: "tango"
    logo: "lacrosse_crop_square.jpg"
    favicon: "lacrosse_crop_square_favicon.png"
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(htmltools)
```



### Interactive plot of congressional word usage by party built with plotly {data-commentary-width=700}

```{r wordfreq}
library(httr)
library(jsonlite)
library(tidyverse)
library(plotly)

apikey <- read_file('~/Documents/R projects/sunlight_api_key.txt')
phrase <- c('rich', 'middle class', 'poor')

wordfreq <- crossing(phrase = factor(phrase, levels = phrase), 
                     party = c('D', 'R')) %>% 
  rowwise() %>%
  mutate(url = paste0('http://www.capitolwords.org/api/1/dates.json?phrase=', 
              phrase, '&party=', party, 
              '&percentages=true&granularity=year&apikey=', apikey),
         url = URLencode(url)) %>% 
  mutate(response = list(GET(url)),
         response = content(response, as = 'text'),
         response = fromJSON(response)) %>%
  unnest() %>% 
  mutate(year = as.integer(year))

# calculate subplot heights
heights <- prop.table(tapply(wordfreq$count, wordfreq$phrase, max))

wordfreq %>% split(.$phrase) %>% 
  map(plot_ly, type = 'bar', x = ~year, y = ~count, color = ~party, 
      colors = c('#4575b4', '#d73027')) %>% 
  subplot(nrows = 3, shareX = TRUE, titleX = FALSE, titleY = TRUE, 
          heights = c(.166, .4, .434)) %>% 
  layout(title = 'Congressional Word Usage by Party',
         xaxis = list(domain = c(0.1, 1)), 
         yaxis = list(title = 'rich', dtick = 500, rangemode = 'tozero'),
         yaxis2 = list(title = 'middle class'), 
         yaxis3 = list(title = 'poor'),
         showlegend = FALSE, 
         updatemenus = list(list(yanchor = 'top',
                                 y = 1,
                                 buttons = list(
                                   list(args = list("type", "bar"),
                                        label = "Bar"),
                                   list(args = list("type", "line"),
                                        label = "Line"))),
                            list(yanchor = 'top',
                                 y = .8,
                                 buttons = list(
                                   list(args = list('line', list(shape = 'linear')),
                                        label = 'Linear'),
                                   list(args = list('line', list(shape = 'hvh')),
                                        label = 'Step'),
                                   list(args = list('line', list(shape = 'spline')),
                                        label = 'Spline')))))
```

***

```r
library(tidyverse)
library(plotly)

# data from Sunlight Foundation Capitol Words API

wordfreq %>% split(.$phrase) %>% 
    map(plot_ly, type = 'bar', x = ~year, y = ~count, color = ~party, 
        colors = c('#4575b4', '#d73027')) %>% 
    subplot(nrows = 3, shareX = TRUE, titleX = FALSE, titleY = TRUE, 
            heights = c(.166, .4, .434)) %>% 
    layout(title = 'Congressional Word Usage by Party',
           xaxis = list(domain = c(0.1, 1)), 
           yaxis = list(title = 'rich', dtick = 500, rangemode = 'tozero'),
           yaxis2 = list(title = 'middle class'), 
           yaxis3 = list(title = 'poor'),
           showlegend = FALSE, 
           updatemenus = list(
               list(yanchor = 'top', y = 1,
                    buttons = list(
                        list(args = list("type", "bar"),
                             label = "Bar"),
                        list(args = list("type", "line"),
                             label = "Line"))),
               list(yanchor = 'top', y = 0.8,
                    buttons = list(
                        list(args = list('line', list(shape = 'linear')),
                             label = 'Linear'),
                        list(args = list('line', list(shape = 'hvh')),
                             label = 'Step'),
                        list(args = list('line', list(shape = 'spline')),
                             label = 'Spline')))))
```


### Animation of DC home values built with ggplot2 and gganimate {data-commentary-width=700}

```{r zillow}
img(src = image_uri('~/Documents/R projects/mapp/dc_zip_zhvi-lolli.gif'))
```

***

```r
library(tidyverse)
library(gganimate)

# data from Zillow
house <- read.csv('Zip_Zhvi_AllHomes.csv')
dczip <- house %>% filter(State == 'DC') %>% 
    gather(month, zhvi, starts_with('X')) %>% 
    filter(!is.na(zhvi))

plots <- ggplot(dczip, 
                aes(x = RegionName, y = zhvi, colour = zhvi, frame = month), 
                show.legend = F) + 
    geom_point() + 
    geom_segment(aes(xend = ..x.., yend = 0)) + 
    scale_colour_gradient(low = '#2171b5', high = '#e34a33') + 
    labs(x = 'Zip Code', y = 'Zillow Home Value Index', title = 'Washington DC Home Values') + 
    scale_y_continuous(labels = scales::comma) + 
    theme(text = element_text(family = "Myriad Pro", colour = "gray30"),
          axis.title = element_text(size = 9, face = 'bold'),
          plot.title = element_text(face = 'bold'), 
          axis.ticks = element_line(colour = 'gray70'),
          legend.position = "none")

gg_animate(plots, saver = 'gif',
           interval = 0.02, ani.type = 'svg', 
           ani.width = 600, ani.height = 400)
```

### Leaflet map of candidates running for president from FEC data {data-commentary-width=700}

```{r leaflet}
library(dplyr)
library(leaflet)

candidates <- read.csv('~/Documents/R projects/running_for_president/candidates.csv')

# build HTML for leaflet marker popups and pare unnecessary columns
candidates_for_map <- candidates %>% filter(!is.na(lat)) %>% 
    mutate(popup = paste(
        '<table><tr><th>Candidate:</th><th>', can_nam, '</th></tr>',
        '<tr><td>Address:</td><td>', 
        can_str1, can_str2, '<br />',
        can_cit, can_sta, can_zip, '</td></tr>',
        '<tr><td>Net Contributions:</td><td>', net_con,
        '</td></tr></table>'
    )) %>% 
    select(lat, lng = lon, popup)

leaflet(candidates_for_map) %>% 
  setView(-95.71289, 37.09024, zoom = 4) %>%
  addProviderTiles('CartoDB.Positron') %>% 
  addMarkers(
    lat = ~lat, 
    lng = ~lng,
    popup = ~popup, 
    clusterOptions = markerClusterOptions()
  )
```

***

```r
library(dplyr)
library(leaflet)

# candidate data from http://fec.gov/data/CandidateSummary.do?format=html
candidates <- read.csv('candidates.csv')

# build HTML for leaflet marker popups
candidates_for_map <- candidates %>% filter(!is.na(lat)) %>% 
    mutate(popup = paste(
        '<table><tr><th>Candidate:</th><th>', can_nam, '</th></tr>',
        '<tr><td>Address:</td><td>', 
        can_str1, can_str2, '<br />',
        can_cit, can_sta, can_zip, '</td></tr>',
        '<tr><td>Net Contributions:</td><td>', net_con,
        '</td></tr></table>'
    )) %>% 
    select(lat, lng = lon, popup)

leaflet(candidates_for_map) %>% 
    setView(-95.71289, 37.09024, zoom = 4) %>%
    addProviderTiles('CartoDB.Positron') %>% 
    addMarkers(
        lat = ~lat, 
        lng = ~lng,
        popup = ~popup, 
        clusterOptions = markerClusterOptions()
    )
```


### Trump presidential campaign disbursement sunburst {data-commentary-width=700}

```{r sunburst}
library(readr)
library(sunburstR)

# disbursement data from http://fec.gov/data/CandidateSummary.do?format=html
donald <- read_csv('~/Documents/R projects/fec_disbursements/Two_Year_Summary_Operating_Expenditures.csv',
                   skip = 7, trim_ws = TRUE,
                   col_types = cols(`Payment Date` = col_date(format = '%m/%d/%Y'),
                                    Amount = col_number()))

donald %>% mutate(Category = gsub('^(\\w+).*', '\\1', Purpose)) %>% 
  mutate_at(c('Purpose', 'Category', '`Payee Name`'), funs(gsub('-', ' ', .))) %>% 
  mutate(Categories = paste(Category, 
                            Purpose, 
                            `Payee Name`, 
                            `Report Year`, 
                            `Report Type`, sep = '-')) %>% 
  select(Categories, Amount) %>% 
  sunburst()
```

***

```r
library(tidyverse)
library(sunburstR)

# disbursement data from http://fec.gov/data/CandidateSummary.do?format=html
donald <- read_csv('Two_Year_Summary_Operating_Expenditures.csv',
                   skip = 7, trim_ws = TRUE,
                   col_types = cols(`Payment Date` = col_date(format = '%m/%d/%Y'),
                                    Amount = col_number()))

donald %>% mutate(Category = gsub('^(\\w+).*', '\\1', Purpose)) %>% 
    mutate_at(c('Purpose', 'Category', '`Payee Name`'), funs(gsub('-', ' ', .))) %>% 
    mutate(Categories = paste(Category, 
                              Purpose, 
                              `Payee Name`, 
                              `Report Year`, 
                              `Report Type`, sep = '-')) %>% 
    select(Categories, Amount) %>% 
    sunburst()
```

### About Me {style="padding: 0px 10%;"}

```{r about_me}
div(includeMarkdown('about_me.md'), style = 'padding: 25px 10%;')
```

***

```{r contact}
div(includeMarkdown('about_me_sidebar.md'), style = 'padding: 25px 10%;')
```
